\documentclass[etd,oneside,phd]{BYUPhys}

\usepackage{amsmath}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{float}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[caption=false]{subfig}
\usepackage{pifont}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{mathtools}
\usepackage{microtype}
\usepackage{tikz}
\usepackage{caption}
\usepackage{booktabs}
% Better page breaking / avoid widows and orphans
\usepackage[all]{nowidow}
\clubpenalty=10000
\widowpenalty=10000
\displaywidowpenalty=10000

% Control figure placement
\usepackage{placeins}
\everymath{\displaystyle}

\newcommand{\cmark}{{\color{green!60!black}\ding{51}}}
\newcommand{\xmark}{{\color{red!70!black}\ding{55}}}
% If you'd like to print your thesis using double-sided printing,
% remove the etd option and change the oneside argument to twoside, like this:
%
%\documentclass[twoside,masters]{BYUPhys}
\newcommand{\fignote}[1]{%
  \caption*{\textit{Note.} #1}%
}

\newcommand{\tabnote}[1]{%
  \caption*{\textit{Note.} #1}%
}
% Your name
  \Author{Andrew James Carroll}

% Enter the year your thesis is approved (for senior thesis or capstone)
  \Year{[2026]}

% If you have a long title, split it between multiple lines using the \\ command
  \Title{Numerical Relativity Beyond General Relativity: Einstein Maxwell Dilaton Axion Black Holes and Compact Finite Differences}

% Your research advisor
\AdvisorTitle{Advisors}
  \Advisor{Eric Hirschmann}

% The text of your abstract
\Abstract{This dissertation analyzes two complementary aspects of numerical relativity.  First, we generalize our numerical relativity code to evolve black holes in alternative theories of gravity.  In particular, we develop a numerical framework for studying black holes in Einstein--Maxwell--Dilaton--Axion theory.  Black holes in this theory can support additional scalar and pseudoscalar fields, or hair, that are not present for black holes in general relativity.  We present a new method for generating approximate binary black-hole initial data in EMDA theory and use head-on collisions to demonstrate the stability of charged black holes in this setting.  We then develop a modest catalog of charged black-hole waveforms and compare these waveforms with general-relativistic reference waveforms.  These comparisons show that the additional fields can produce changes in the gravitational waveform, with differences that generally increase with charge. We further explore the broader theory space within the EMDA theory. 

Second, we study compact finite-difference derivative operators for use in adaptive numerical simulations.  These operators are of interest because they have the potential to improve the accuracy and efficiency of numerical relativity calculations.  We construct several novel compact finite-difference operators and test their performance on a hierarchy of problems, including the Maxwell equations, the nonlinear sigma model, Teukolsky-wave evolutions, and evolutions using the BSSN equations.  These tests allow us to compare convergence, stability, and robustness across both linear and nonlinear systems.  We show that compact finite-difference operators can improve performance compared to their explicit counterparts, while also identifying the importance of stability, dissipation, and refinement-boundary behavior.}

\Keywords{Numerical relativity, alternative theories of gravity, black holes, gravitational waves, compact finite-difference methods, adaptive mesh refinement}

% Acknowledge those who helped and supported you
  \Acknowledgments{
  }

%% The members of your committee (masters only need A and B, PhD need all 4)
 \MemberA{David Neilsen}
 \MemberB{Gus Hart}
 \MemberC{Chris Verharen}
 \MemberD{Ben Frandsen}


\begin{document}

 % Start page counting in roman numerals
 \frontmatter

 % This command makes the formal preliminary pages.
 % You can comment it out during the drafting process if you want to save paper.
 \makepreliminarypages

 % Make the table of contents.
 \tableofcontents
\cleardoublepage
\listoffigures

\cleardoublepage
\listoftables
 % Start regular page counting at page 1
 \mainmatter

% OK. Everything is set up. Type your thesis here.
\chapter{Introduction}
\include{Introduction}
\chapter{EMDA Chapter 1}
\include{EMDA_Chapter_1}
\chapter{EMDA Chapter 2}
\include{EMDA_Chapter_2}
\chapter{EMDA Chapter 3}
\include{EMDA_Chapter_3}
\chapter{CFD Chapter 1}
\include{CFD_Chapter_1}
\chapter{Conclusion}
\include{Conclusion}
\cleardoublepage
\begin{appendices}

\chapter{Mathematical Overview}
\include{Mathematical_overview}

\chapter{EMDA Black Holes}
\include{EMDA_Black_holes}

\chapter{EMDA Initital data}
\include{EMDA_Initital_data}

\chapter{Computation of \texorpdfstring{$\Psi_4$}{Psi4}}
\include{appendix_psi_4}

\chapter{Dendro Notes}
\include{Dendro_notes}

\chapter{CFD Notes}
\include{CFD_notes}

\include{Software_notes}

\include{Paraview}

\chapter{Waveform Catalog}
\include{Waveform_catalog}

\end{appendices}


% Make the bibliography.
\bibliography{references}

% Make the index
\printindex

% Make the index
 \printindex

\end{document}
