\section{Introduction}
Recent improvements in gravitational-wave detectors have opened a new regime for precision tests of general relativity (GR). Current observations already reach signal-to-noise ratios high enough to place increasingly stringent constraints on deviations from GR . As detector sensitivity continues to improve, gravitational waves will provide an increasingly powerful probe of strong-field gravity, where compact objects probe highly dynamical and nonlinear spacetime curvature.

This motivates the study of well-posed alternative theories of gravity and their predictions for compact-binary dynamics. A number of theories beyond GR have been considered in this context, including scalar-tensor theories, dynamical Chern--Simons gravity, and Einstein--Gauss--Bonnet gravity. In several of these theories, black holes can acquire additional structure or ``hair,'' leading to modified inspiral dynamics, merger behavior, or gravitational-wave emission. However, despite this progress, comparatively less work has been done on fully dynamical binary black hole evolutions in Einstein--Maxwell--dilaton--axion (EMDA) theory.

EMDA theory is a particularly interesting target for numerical study for several reasons. It provides a well-posed extension of GR coupled to electromagnetic, dilaton, and axion fields, and it arises as a low-energy limit of heterotic string theory. The theory also admits rotating charged black-hole solutions, such as the Kerr--Sen spacetime, which differ from Kerr black holes while remaining observationally viable in some regimes. Previous work has investigated the stability of black-hole hair in EMDA theory \cite{PopeRohrerWhiting2025}, and black-hole shadow studies using Event Horizon Telescope observations have not ruled out Kerr--Sen black holes.

These facts motivate a systematic study of binary black hole dynamics in EMDA theory. In this work, we investigate binary inspirals in EMDA and examine how the additional fields modify the dynamics and gravitational-wave signal relative to GR. Our goal is to provide a step toward connecting EMDA theory with future strong-field gravitational-wave tests.

The cases we consider here are more highly charged than are astrophysically likely. There have been a few mechanisms proposed by which black holes may obtain a charge but any charges are likely to be quite small. We have previously established the stability of EMDA black holes. 

% -------------\subsection{Motivation}

In this work, we present a modest catalog of binary black-hole waveforms in Einstein--Maxwell--dilaton--axion (EMDA) theory, considering several choices of mass ratio, spin, and charge. The parameter space of compact-binary waveforms is highly degenerate. Consequently, it is not always clear whether deviations from a GR waveform arise from genuinely new physics or whether they could instead be mimicked by changes in the intrinsic binary parameters, such as the mass ratio, spins, charges, or total mass scale.

A complete study of these degeneracies would require a full parameter-estimation analysis over the EMDA and GR parameter spaces, which is beyond the scope of this work. Instead, we consider a more restricted question. We assume that the spins are independently constrained and compare EMDA waveforms with corresponding GR waveforms while allowing for a rescaling of the total mass. This comparison allows us to test whether the dominant differences between EMDA and GR waveforms can be absorbed into a simple mass rescaling, or whether theory-dependent residuals remain.

\subsection{Goal of the Work}

The primary goal of this work is to determine under what circumstances EMDA waveforms differ noticeably from their GR counterparts. We focus on how the additional EMDA fields and charges modify the inspiral dynamics and gravitational-wave signal. By comparing EMDA and GR waveforms after accounting for possible mass rescalings, we identify cases in which the waveform differences remain distinguishable.

The waveform catalog presented here is intended as an initial step toward a more complete study of EMDA binary black-hole phenomenology. While we do not attempt to cover the full parameter space or perform a complete parameter-estimation analysis, these results provide useful guidance for future studies aimed at determining whether high-precision gravitational-wave observations can constrain or detect signatures of EMDA theory.

\subsection{Structure of the Paper and Conventions}

The remainder of this paper is organized as follows. In Sec.~\ref{sec:methods}, we review the EMDA equations, describe the numerical methods used for the evolutions, and discuss the construction of initial data. We then describe the procedure used to reduce orbital eccentricity and the waveform-extraction methods used in the analysis. In Sec.~\ref{sec:results}, we compare EMDA and GR waveforms and identify the circumstances under which their differences may become noticeable. We summarize our conclusions and discuss future directions in Sec.~\ref{sec:conclusion}.

Unless otherwise stated, we use geometrized units with $G=c=1$. Greek indices denote spacetime components, while Latin indices denote spatial components.

\section{Methods}
\label{sec:methods}

\section{Einstein--Maxwell--Dilaton--Axion Theory}
\label{sec:emda}

The action for EMDA theory is
\begin{equation}
\begin{aligned}
S=\int \mathrm{d}^4 x\,\sqrt{-g}\Big[
&R-2(\nabla\phi)^2-e^{-2\alpha_0\phi}F^2\\
&-\frac{1}{2} e^{4\alpha_1\phi}(\nabla\kappa)^2-\kappa\,F_{ab}({}^*F)^{ab}
\Big].
\end{aligned}
\end{equation}
Here, $g$ denotes the determinant of the spacetime metric, $R$ is the Ricci scalar, $\phi$ is the dilaton field, $F_{ab}$ is the Faraday tensor, and $\kappa$ is the axion field. The parameters $\alpha_0$ and $\alpha_1$ control the coupling strengths of the dilaton and axion sectors. In this work, we take $\alpha_0=\alpha_1=1$, corresponding to the low-energy limit of heterotic string theory. Setting $\alpha_0=\alpha_1=0$ reduces the theory to Einstein--Maxwell theory with minimally coupled scalar and axion fields.

In the absence of charge, the electromagnetic field vanishes and the equations reduce to the standard Einstein equations. When charge is present, however, the electromagnetic field sources the dilaton and axion fields. In this sense, EMDA black holes possess additional field structure beyond the Kerr family. These extra fields provide a possible source of deviations from GR in compact-binary dynamics and gravitational-wave emission.

To perform numerical evolutions of Kerr--Sen black holes, we formulate the equations as a Cauchy problem and use the BSSN formulation \cite{Shibata1995_BSSN, Baumgarte1999_BSSN, Baumgarte2010} of the Einstein equations, supplemented by the electromagnetic, dilaton, and axion source terms. We also employ the moving-puncture gauge conditions \cite{Campanelli2006,Baker2006}. The full equations of motion and their $3+1$ decomposition are presented in our earlier paper.

\subsection{Controlling the Eccentricity}

We use the same \texttt{TPID} solver described in our earlier work to construct binary black-hole initial data in EMDA theory. To set approximately quasicircular initial data, we use a post-Newtonian prescription implemented in the script developed by Zachariah Etienne. In applying this procedure, we assume that the direct contributions from the dilaton and axion fields to the initial orbital parameters are negligible.

We then follow the procedure used for quasicircular charged black-hole initial data in Ref.~\cite{Bozzola2021}. Although this approach is approximate, it produces relatively low eccentricity for the spinless configurations considered here. After excluding the first orbit, we find eccentricities of approximately \textbf{[insert eccentricities here]}.

\subsection{Wave Extraction}

To analyze the numerical data, we compute the Newman--Penrose scalar $\Psi_4$ and extract its multipole moments at finite radius. We also compute the corresponding radiated electromagnetic quantities. Unless otherwise stated, waveform quantities are extracted at $r \approx 100M$. Further details of the waveform-extraction procedure are given in Appendix~\ref{app:wave-extraction}.

\subsection{Code Description}

For these simulations, we modified the existing numerical-relativity framework \texttt{Dendro-GR}. \texttt{Dendro-GR} is a portable code that combines an octree-based adaptive multiresolution framework with automatic code generation for solving relativistic field equations. In this work, we extend the code to evolve the EMDA equations.

Spatial derivatives are computed using sixth-order finite-difference stencils, and time integration is performed using a fourth-order Runge--Kutta scheme. Additional details on \texttt{Dendro-GR} and its numerical infrastructure can be found in Refs.~\cite{Fernando2019,Fernando2019b,Fernando2023,Black2025_ORBIT}.

\subsection{Simulation Setup}

A summary of the black-hole parameters used in the simulations is provided in Appendix~\ref{app:catalog}. The catalog includes configurations with different mass ratios, spins, and charges. These simulations are designed to sample representative regions of the EMDA parameter space rather than to provide a complete parameter survey.

\section{Results}
\label{sec:results}

We compare EMDA and GR waveforms for a number of representative configurations. In particular, we examine how the waveform differences depend on charge, mass ratio, and spin. We also compute signal-to-noise ratios and compare strains after allowing for an overall mass rescaling. This allows us to assess whether EMDA waveforms can be effectively mimicked by nearby GR configurations or whether residual differences remain.

Our results suggest that, for astrophysically plausible charges, distinguishing EMDA waveforms from GR waveforms is likely to be challenging with current detector sensitivities. Larger charges can produce more noticeable deviations, but such configurations may be less astrophysically likely.

\subsection{SNR and Strain Comparison}

\subsection{Detectability}

\section{Conclusion}
\label{sec:conclusion}

In this work, we presented a modest catalog of binary black-hole waveforms in EMDA theory. We compared these waveforms with corresponding GR waveforms and examined whether the leading differences could be absorbed by a rescaling of the total mass. This provides an initial test of whether EMDA corrections produce distinguishable signatures in gravitational-wave observations.

We find that waveform differences are generally small for astrophysically plausible charges, suggesting that EMDA effects may be difficult to distinguish from GR in such regimes. However, for larger charges, the deviations become more pronounced and may provide useful targets for future studies. A more complete parameter-estimation analysis will be needed to determine whether these effects can be robustly distinguished from changes in mass ratio, spin, or other binary parameters.

Future work should extend this catalog to a broader region of the EMDA parameter space, include longer inspirals and higher-resolution simulations, and perform direct comparisons with detector noise curves and Bayesian parameter-estimation pipelines.

\appendix

\section{Waveform Catalog}