\section{Initial Data}
\label{sec:appx:id}

It is necessary to have good initial data for our EMDA BH binaries.  Of course, exact solutions are not known.  But we can generalize standard approaches for the construction of initial data in vacuum GR to these additional fields on making certain approximations.  We follow closely the approach based on the conformal transverse-traceless decomposition  used with punctures.  Indeed, we extend this approach as it has been incorporated into the {\tt TwoChargedPunctures} code of Bozzola and Paschalidis~\cite{Bozzola2019,ferreira2026electromagneticdualitydegeneracydynamical}.  We mostly follow their notation below.

As usual, we decompose the metric, but using notation slightly modified from the previous appendix.  In particular we define 
%systemsAs a first step, we assume conformally flat space; that is, we decompose the metric as
\begin{equation}
    \gamma_{ij}= \psi^4 \bar{\gamma}_{ij},
\end{equation}
with $\psi$ the conformal factor and the conformal metric assumed to be the flat metric, ${\bar\gamma}_{ij} = \eta_{ij}$.  The latter assumption simplifies the following computations and is a reasonable choice for the systems we study. It is a limiting assumption preventing the construction of highly spinning black holes~\cite{ValienteKroon2004,Price2000}. 
Notwithstanding, Bowen--York black holes can reach spin values of order $\chi \sim 0.9$ \cite{Dain2002}. As our simulations remain below this limit, the approximation should be reasonable.  
While we anticipate the appearance of ``junk radiation'' partly in consequence of this assumption~\cite{Bozzola2021}, we find that it does decay  
as the fields relax to quasi-equilibrium which results in decreased constraint violation over the course of the evolutions. We adopt the maximal slicing condition, $K=0$, and define a conformally rescaled tracefree extrinsic curvature, ${\bar A}_{ij}=\psi^{2}(K_{ij} - \tfrac{1}{3} \gamma_{ij} K)$.  Ignoring the latter's radiative degrees of freedom, we can express it in terms of a vector, $V^i$, as
\begin{equation}
    \bar{A}_{ij} = \partial_i V_j + \partial_j V_i - \frac{2}{3}\eta_{ij}\partial_k V^k.
\end{equation} 

The initial data problem then reduces to determining $V^i$ and $\psi$ for the gravitational pieces.  In particular, the Hamiltonian and momentum constraint equations take the form
\begin{equation}     
\partial_k\partial^k \psi + \tfrac{1}{8} \psi^{-7} \bar{A}_{ij} \bar{A}^{ij} = - 2\pi \psi^5 \rho , \\
\end{equation}
\label{hamiltonian_constraint} 
\begin{equation}
    \partial_k \partial^k V^i + \tfrac{1}{3}\eta^{ij}\partial_j(\partial_k V^k ) = 8 \pi S^i . 
\label{momemtum_constraint}
\end{equation}

The momentum constraint is linear and independent of $\psi$.  This permits dividing the solution between homogeneous (vacuum gravitational) and inhomogeneous (matter) contributions:
\begin{equation}
    V^i = V^i_{\text{GR}} + V^i_{\text{M}}.
\end{equation}
For $V^i_{\text{GR}}$, we use the standard Bowen--York solutions~\cite{Bowen1980} while for the matter contribution, we must solve Eq.~\ref{momemtum_constraint} with 
\begin{equation}
    S^i = 2 \, \Pi \, D^i \phi 
    + 2 e^{-2 \alpha_0 \phi} \epsilon^{ijk} E_j B_k 
    + \tfrac{1}{2}e^{4 \alpha_1 \phi} \, \Xi \, D^i \kappa.
\end{equation}
If we assume a moment of time symmetry and neglect contributions from the magnetic field this source term vanishes.  On imposing asymptotically flat boundary conditions, the solution to the matter part therefore becomes $V^i_{\text{M}}=0$.  

We must also impose the electromagnetic constraints. At this point we  depart somewhat from~\cite{Bozzola2019} because the EMDA version of these constraints are no longer linear, containing rather a nontrivial current that couples the dilaton and axion to the electromagnetic fields: 
\begin{align}
\nabla_i E^i &= 2\alpha_0 E^i \partial_i \phi + e^{2\alpha_0 \phi} B^i \partial_i \kappa \\
\nabla_i B^i &= 0 
\end{align} 
Were the source in the Gauss constraint zero, we would 
follow \cite{Alcubierre2009,Zilhao2012} and perform a conformal rescaling of the fields as
\begin{equation}
    \bar{E}^i = \psi^6 E^i, 
    \quad
    \bar{B}^i = \psi^6 B^i.  
\end{equation}
This results in the sourceless Maxwell constraints taking the form 
\begin{equation}
    \partial_i \bar{E}^i = 0, 
    \quad
    \partial_i \bar{B}^i = 0,
\end{equation}
independent of $\psi$ and therefore solvable separately from the spacetime variables. Since these equations are linear, solutions may be superposed. The advantages of this are significant so we choose to make the approximation that this source term is, in fact, negligible and we solve the EMDA electromagnetic constraints in the way just described.  Of course, this failure to self-consistently solve these constraints necessarily introduces violations.  Partly for this reason, we have introduced hyperbolic divergence cleaning to drive these violations smaller over the course of the evolution.  We have measured these violations on the initial slice. For the simulations considered in this work we confirm that they are consistently small and driven smaller via the constraint damping. The expected damping behavior is demonstrated in Figure~\ref{fig:constraint_divb}. There is an initial violation of the EM constraints, which damps over time.    

The Hamiltonian constraint above, not being linear, must be treated differently.  We assume that we have two black holes charged under the U(1) gauge field as well as possessing scalar dilaton and axion fields.  Following \cite{Bozzola2019}, we approximate this with a conformal factor of the form 
\begin{equation}
    \psi^2 = (1 + u + \bar{\eta} )^2 - \bar{\xi}^2 
\end{equation}
for which we define 
\begin{align}
   \bar{\eta} &= \frac{m_1}{2R_1} + \frac{m_2}{2R_2}, \\
   \bar{\xi}  &= \frac{q_1}{2R_1} + \frac{q_2}{2R_2}, \\ 
   \bar{\kappa} &= 1+ u + {\bar\eta},
\end{align}
and where $m_i$, $q_i$ and $R_i$ here are the masses, electric charges and the coordinate distances from the origin of the BHs.  This is intended to correspond to two electrically charged BHs with the singular part of the conformal factor separated out with $u$ giving finite corrections, including the necessary corrections for the additional scalar fields. 

On rearranging the Hamiltonian constraint in terms of $u$, we find 
\begin{multline}
    \partial_k\partial^k u = -\frac{1}{\bar{\kappa}} \Big(
    \partial_a \bar{\kappa} \partial^a \bar{\kappa} 
    - \partial_a \bar{\xi} \partial^a \bar{\xi} 
    - \partial_a \psi \partial^a \psi \\
    + \tfrac{1}{8} \psi^{-6} \bar{A}_{ij} \bar{A}^{ij}
    + 2 \pi \psi^{6} \rho
    \Big),
\end{multline}
with
\begin{equation}
\begin{split}
    \rho \equiv\;& \bigl[ \Pi^2 + (D\phi)^2 \bigr]
    + e^{-2 \alpha_0 \phi}\big[ E_a E^a +  B_a  B^a\big] \\
    &+ \tfrac{1}{4} e^{4 \alpha_1 \phi}\big[\Xi^2 + (D\kappa)^2\big].
\end{split}
\end{equation}
Using simplifying assumptions similar to above, namely a moment of time symmetry and neglecting the magnetic field, $\rho$ reduces to
\begin{equation}
    \rho = (D\phi)^2 + e^{-2 \alpha_0 \phi} \, E_a E^a + \tfrac{1}{4} e^{4\alpha_1 \phi} \, (D\kappa)^2 .
\end{equation}
Again, this follows \cite{Bozzola2019} quite closely, except that we have now included dilaton and axion contributions in $\rho$. Specifically, we employ a Reissner--Nordström electric field for the electric component and add a dilaton field for a single, spinless black hole. In practice, the axion is initialized to zero and allowed to relax to a quasi-equilibrium state during an evolution.  

The electric field is initially set to be:
\begin{equation}
     E^i = \psi^{-6} \Bigl[ \frac{q_1}{R_1^2} + \frac{q_2}{R_2^2} \Bigr].
\end{equation}
while the dilaton fields are initially set as:
\begin{align}
\phi &= \phi_1(r_1,\theta_1) + \phi_2(r_2, \theta_2) 
\end{align}
where Eq.~\ref{Kerr-Sen_dilaton} gives the form for a dilaton field for a single Kerr--Sen black hole (albeit in Boyer--Lindquist type coordinates) for our current theory parameters of $(\alpha_0,\alpha_1) = (1,1)$. 
