\section{Electricity and Magnetism}

We give here an example of how to perform a \(3+1\) decomposition for electricity and magnetism. We follow an approach similar to that of \textit{Numerical Relativity from Scratch}~\cite{BaumgarteShapiro2021}.

We begin with Maxwell's equations written in terms of the Faraday tensor:
\begin{equation}
    \nabla_b F^{ab} = 4\pi j^a ,
    \label{eq:maxwell_equation}
\end{equation}
and
\begin{equation}
    \nabla_{[a}F_{bc]}=0 ,
\end{equation}
where we have defined the four-current as
\begin{equation}
    j^a = (\rho, j^i).
\end{equation}
We note that the continuity equation can be written as
\begin{equation}
    \nabla_a j^a = 0 .
\end{equation}
In general relativity, there is a similar conservation law associated with the Bianchi identity.

We further note that the Maxwell tensor can be written in terms of the vector potential \(A^a\) as
\begin{equation}
    F^{ab} = \nabla^a A^b - \nabla^b A^a .
\end{equation}
With this definition, the second of Maxwell's equations is satisfied automatically. Substituting this expression into the first Maxwell equation, Eq.~\eqref{eq:maxwell_equation}, gives
\begin{equation}
    \nabla_b \nabla^a A^b - \nabla_b \nabla^b A^a = 4\pi j^a .
    \label{eq:maxwell_vector_potential}
\end{equation}
Here \(A^a\) is the vector potential. In a manner similar to the four-current, we write the vector potential as
\begin{equation}
    A^a = (\phi, A^i).
\end{equation}

We now first consider the time component of Eq.~\eqref{eq:maxwell_vector_potential}. Setting the index \(a\) to the time component, \(t\), we have
\begin{equation}
    \nabla_b \nabla^t A^b - \nabla_b \nabla^b A^t
    =
    \nabla_b \nabla^t A^b - \nabla_j \nabla^j \phi
    =
    4 \pi j^t
    =
    4 \pi \rho .
\end{equation}
Here we have changed the index from \(b\) to \(j\) because the remaining components are purely spatial. We also note that, when \(b=t\), the two terms cancel. In special relativity, corresponding to flat spacetime, we make the following identifications:
\begin{equation}
    \nabla_j = D_j,
    \qquad
    \nabla_t = \partial_t .
\end{equation}
We also have \(\nabla^t = \eta^{ta}\nabla_a = -\partial_t\), and therefore
\begin{equation}
    D_j \left( \partial_t A^j - D^j \phi \right) = 4 \pi \rho .
\end{equation}
We can now make the identification
\begin{equation}
    E^i = F^{ti} = -\partial_t A^i - D^i \phi .
\end{equation}
This can also be rearranged to obtain the evolution equation
\begin{equation}
    \partial_t A^i = -E^i - D^i \phi .
\end{equation}
As a result, we recover Gauss's law,
\begin{equation}
    D_i E^i = 4 \pi \rho ,
\end{equation}
which gives the constraint equation.

We now consider the spatial components of Eq.~\eqref{eq:maxwell_vector_potential}, corresponding to \(a=i\):
\begin{equation}
    \nabla_t \left(\nabla^i A^t - \nabla^t A^i \right)
    =
    - \nabla_j \nabla^i A^j
    + \nabla_j \nabla^j A^i
    + 4 \pi j^i .
\end{equation}
Here we have grouped the terms with \(b=t\) on the left-hand side and the terms with \(b=j\) on the right-hand side. We then obtain
\begin{equation}
    \partial_t E_i
    =
    - D_j D^j A_i
    + D_i D^j A_j
    - 4 \pi j_i .
\end{equation}
We now have both the evolution and constraint equations. We will follow the same pattern as in general relativity, but we will have to contend with non-flat spacetime.

To summarize, we have six evolution equations:
\begin{equation}
    \partial_t E_i
    =
    - D_j D^j A_i
    + D_i D^j A_j
    - 4 \pi j_i .
\end{equation}
\begin{equation}
    \partial_t A^i = -E^i - D^i \phi .
\end{equation}
and two constraint equations:
\begin{equation}
    D_i E^i = 4 \pi \rho ,
\end{equation}
\begin{equation}
    D_i B^i = 0 .
\end{equation}
\section{Spacetime and Kinematics}

We now allow for more general slices of spacetime. This increased freedom is needed for solving the Cauchy problem in general relativity and in alternative theories of gravity. The surfaces we embed in spacetime can be thought of as hypersurfaces. We assume that these hypersurfaces are spacelike, so that they represent slices of constant time.

We can define the gradient of \(t\) as
\begin{equation}
    \nabla_a t = \partial_a t .
\end{equation}
This quantity is perpendicular to the constant-time hypersurfaces. We then introduce the normal \(n_a\) to the slices as proportional to this gradient:
\begin{equation}
    n_a = -\alpha \nabla_a t .
\end{equation}
We choose \(n_a\) to have unit magnitude:
\begin{equation}
    n_a n^a = g^{ab} n_a n_b = -1 .
\end{equation}
Because the slices are spacelike, the normal four-vector must be timelike, and we therefore normalize it to negative unity. The quantity \(\alpha\) measures the ratio of coordinate time to proper time. In general, \(\alpha\) is a coordinate-dependent quantity.

In this coordinate system,
\begin{equation}
    \nabla_a t = \partial_a t = \frac{\partial t}{\partial x^a},
\end{equation}
so the only nonzero component occurs when \(a=t\). The normal vector then reduces to
\begin{equation}
    n_a = (-\alpha,0,0,0).
\end{equation}
We can write \(\alpha\) in terms of the proper time interval \(d\tau\) as
\begin{equation}
    d\tau = \alpha dt .
\end{equation}
We often call \(\alpha\) the lapse function.

The normalization condition can be satisfied by writing
\begin{equation}
    n^a = \frac{1}{\alpha}(1,\beta^i).
\end{equation}
This holds with our normalization condition for any \(\beta^i\). Viewing \(n^a\) as the four-velocity of a normal observer, \(\beta^i\) measures the extent to which the spatial coordinates change as the normal observer moves through spacetime. This quantity is called the shift vector.

We can also use the normal vector to decompose the metric into spatial and temporal parts. We write
\begin{equation}
    g^a{}_b = \gamma^a{}_b - n^a n_b .
\end{equation}
Here \(\gamma^a{}_b\) is the spatial part of the metric and is called the projection operator.

Changes in the normal vector determine how the spatial hypersurfaces are embedded in spacetime. We can decompose this as
\begin{align}
    \nabla_a n_b
    &=
    g_a{}^c g_b{}^d \nabla_c n_d \nonumber \\
    &=
    \left(\gamma_a{}^c - n^c n_a\right)
    \left(\gamma_b{}^d - n^d n_b\right)
    \nabla_c n_d \nonumber \\
    &=
    \gamma_a{}^c \gamma_b{}^d \nabla_c n_d
    - \gamma_a{}^c n_b n^d \nabla_c n_d
    - n_a n^c \gamma_b{}^d \nabla_c n_d
    + n_a n^c n_b n^d \nabla_c n_d .
\end{align}
The Lie derivative measures how a tensor field changes as it is dragged along the flow generated by a vector field. For a covector $A_d$, the Lie derivative along $n^a$ is
\begin{equation}
    \mathcal{L}_n A_d
    =
    n^c \nabla_c A_d
    +
    A_c \nabla_d n^c .
\end{equation}
Unlike the covariant derivative, the Lie derivative does not require any additional connection structure; it depends only on the vector field that generates the flow.

One important property is that the Lie derivative of a spatial tensor along $\alpha n^a$ is again spatial. Since we want our spatial quantities to remain spatial during the evolution, we decompose the coordinate time-evolution vector $t^a$ into normal and tangential parts:
\begin{equation}
    t^a = \alpha n^a + \beta^a ,
\end{equation}
or equivalently,
\begin{equation}
    \alpha n^a = t^a - \beta^a .
    \label{eq:normal}
\end{equation}
Here $\alpha$ is the lapse function, which controls the rate of proper time elapsed between neighboring slices, and $\beta^a$ is the shift vector, which describes how spatial coordinates move from one slice to the next.

We also define the extrinsic curvature by
\begin{equation}
    K_{ab}
    =
    - \gamma_a{}^c \gamma_b{}^d \nabla_c n_d .
\end{equation}
The minus sign is a convention that we adopt.
Before we move on, we will also define the covariant derivative and the Christoffel symbols.

For the covariant derivative, we need a derivative operator that transforms like a tensor. For scalars, we can use the ordinary partial derivative, but for higher-rank tensors this is not sufficient. The partial derivative only accounts for changes in the components of the tensor, while ignoring changes in the basis vectors. To account for changes in the basis vectors, we introduce the Christoffel symbols, which are not themselves tensors.

For a vector $A^b$, the covariant derivative is defined by
\begin{equation}
    \nabla_a A^b
    =
    \partial_a A^b
    +
    \Gamma^b{}_{ac} A^c .
\end{equation}
Here $\Gamma^b{}_{ac}$ is the Christoffel symbol. For the Levi-Civita connection, which is compatible with the metric and torsion-free, the Christoffel symbols are given by
\begin{equation}
    \Gamma^a{}_{bc}
    =
    \frac{1}{2} g^{ad}
    \left(
        \partial_b g_{dc}
        +
        \partial_c g_{db}
        -
        \partial_d g_{bc}
    \right).
\end{equation}
\section{General Relativity}

To decompose Einstein's equations in a similar way, we need the spacetime metric $g_{ab}$. Recall Einstein's equation:
\begin{equation}
    G_{ab} + \Lambda g_{ab} = 8 \pi T_{ab}.
\end{equation}
At first, this appears to provide ten equations for the ten independent components of $G_{ab}$. However, the Bianchi identity,
\begin{equation}
    \nabla_a G^{ab} = 0,
\end{equation}
must always be satisfied. This plays a role similar to the constraints in electromagnetism. It also tells us that four of Einstein's equations are not independent. We therefore expect four constraint equations in general relativity, along with a fourfold gauge freedom.

We can now decompose both the variables and the equations. As mentioned above, the spacetime metric can be decomposed using the spatial projection operator:
\begin{equation}
  g_{ab}
  =
  g^c{}_a g^d{}_b g_{cd}
  =
  \left(\gamma^c{}_a - n^c n_a\right)
  \left(\gamma^d{}_b - n^d n_b\right) g_{cd}
  =
  \gamma^c{}_a \gamma^d{}_b g_{cd} - n_a n_b .
\end{equation}
The first term on the right-hand side is
\begin{equation}
    \gamma^c{}_a \gamma^d{}_b g_{cd}
    =
    g_{ab} + n_a n_b .
\end{equation}
This defines the spatial metric, which also acts as the projection tensor:
\begin{equation}
    \gamma_{ab} = g_{ab} + n_a n_b .
\end{equation}
The spatial metric performs the same role on each spatial slice that the spacetime metric performs in the full four-dimensional spacetime. However, because $\gamma_{ab}$ is purely spatial, its components normal to the slice vanish.

We can therefore write the inverse spacetime metric as
\begin{equation}
    g^{ab}
    =
    \gamma^{ab} - n^a n^b
    =
    \begin{pmatrix}
        -\alpha^{-2} & \alpha^{-2} \beta^i \\
        \alpha^{-2}\beta^j & \gamma^{ij} - \alpha^{-2}\beta^i\beta^j
    \end{pmatrix}.
\end{equation}
The line element then becomes
\begin{equation}
    ds^2
    =
    g_{ab} dx^a dx^b
    =
    -\alpha^2 dt^2
    +
    \gamma_{ij}
    \left(dx^i + \beta^i dt\right)
    \left(dx^j + \beta^j dt\right).
\end{equation}
The quantity $ds^2$ encodes the notion of distance in spacetime and can be thought of as the spacetime analogue of the Pythagorean theorem.

The extrinsic curvature can also be related to the time evolution of the spatial metric. In particular, one can show that
\begin{equation}
    K_{ab}
    =
    -\frac{1}{2}\mathcal{L}_n \gamma_{ab}.
    \label{eq:extrinsic-curvature-lie}
\end{equation}
Using Eq.~\eqref{eq:normal} and Eq.~\eqref{eq:extrinsic-curvature-lie}, this can be rewritten in terms of the coordinate time derivative as
\begin{equation}
    \partial_t \gamma_{ij}
    =
    -2 \alpha K_{ij}
    +
    \mathcal{L}_{\beta}\gamma_{ij}.
\end{equation}
Thus, the extrinsic curvature measures how the spatial metric changes in time, up to the lapse and shift. The spatial metric $\gamma_{ij}$ and the extrinsic curvature $K_{ij}$ together form the basic dynamical variables in the ADM decomposition.

From here, we need to further decompose the equations into purely spatial quantities. The spatial covariant derivative is defined in the same way as the spacetime covariant derivative, but with the connection associated with the spatial metric $\gamma_{ij}$. For a spatial vector $V^j$, we have
\begin{equation}
    D_i V^j
    =
    \partial_i V^j
    +
    \Gamma^j{}_{ik} V^k .
\end{equation}
Here $\Gamma^i{}_{jk}$ is the Christoffel symbol associated with the spatial metric:
\begin{equation}
    \Gamma^i{}_{jk}
    =
    \frac{1}{2} \gamma^{il}
    \left(
        \partial_j \gamma_{lk}
        +
        \partial_k \gamma_{lj}
        -
        \partial_l \gamma_{jk}
    \right).
\end{equation}

To obtain the spatial decompositions of the Riemann and Ricci tensors, we use the Gauss--Codazzi relations. One useful projection is
The spatial projections of the spacetime curvature are related to the intrinsic curvature of the spatial slice and the extrinsic curvature through the Gauss--Codazzi relations. These equations allow us to rewrite spacetime curvature quantities in terms of purely spatial quantities and the extrinsic curvature.

The Gauss equation is
\begin{equation}
    \gamma^a{}_i \gamma^b{}_j \gamma^c{}_k \gamma^d{}_l
    {}^{(4)}R_{abcd}
    =
    {}^{(3)}R_{ijkl}
    +
    K_{ik}K_{jl}
    -
    K_{il}K_{jk}.
\end{equation}
Here ${}^{(4)}R_{abcd}$ is the four-dimensional spacetime Riemann tensor, while ${}^{(3)}R_{ijkl}$ is the three-dimensional Riemann tensor associated with the spatial metric $\gamma_{ij}$.

The Codazzi equation is
\begin{equation}
    \gamma^a{}_i \gamma^b{}_j \gamma^c{}_k n^d
    {}^{(4)}R_{abcd}
    =
    D_j K_{ik}
    -
    D_i K_{jk}.
\end{equation}
This relates the mixed spatial-normal projection of the spacetime curvature to spatial derivatives of the extrinsic curvature.

The Ricci equation, which involves two projections along the normal direction, is
\begin{equation}
    \gamma^a{}_i \gamma^b{}_j n^c n^d
    {}^{(4)}R_{acbd}
    =
    \mathcal{L}_n K_{ij}
    +
    \frac{1}{\alpha} D_i D_j \alpha
    +
    K_{ik}K^k{}_j .
\end{equation}
Together, these relations provide the geometric bridge between the four-dimensional spacetime curvature and the three-dimensional geometry of each spatial slice.
The preceding definitions come from differential geometry. To obtain the ADM equations, we now insert them into Einstein's equation and equate the projections of the Einstein tensor with the corresponding projections of the stress-energy tensor in Eq.~\eqref{eq:Einstein}.

Projecting Einstein's equation twice along the normal direction gives the Hamiltonian constraint:
\begin{equation}
    R + K^2 - K_{ij}K^{ij}
    =
    16\pi \rho .
    \label{eq:hamiltonian}
\end{equation}
Here
\begin{equation}
    \rho = n^a n^b T_{ab}
\end{equation}
is the energy density measured by an observer moving normal to the spatial slice. The quantity $K$ is the trace of the extrinsic curvature,
\begin{equation}
    K = \gamma^{ij}K_{ij}.
\end{equation}

A mixed projection, with one index projected along the normal direction and one projected onto the spatial slice, gives the momentum constraint:
\begin{equation}
    D_j \left( K^{ij} - \gamma^{ij}K \right)
    =
    8\pi j^i .
    \label{eq:momentum}
\end{equation}
Here
\begin{equation}
    j^i
    =
    - \gamma^{ia} n^b T_{ab}
\end{equation}
is the momentum density measured by the normal observer.

A complete spatial projection gives the evolution equation for the extrinsic curvature:
\begin{equation}
    \partial_t K_{ij}
    =
    \alpha
    \left(
        R_{ij}
        - 2K_{ik}K^k{}_j
        + K K_{ij}
    \right)
    - D_i D_j \alpha
    - 4\pi \alpha M_{ij}
    + \mathcal{L}_{\beta} K_{ij}.
    \label{eq:adm-k-evolution}
\end{equation}
In this expression, we have used the abbreviation
\begin{equation}
    M_{ij}
    =
    2S_{ij}
    -
    \gamma_{ij}\left(S-\rho\right),
\end{equation}
where
\begin{equation}
    S_{ij}
    =
    \gamma^a{}_i \gamma^b{}_j T_{ab}
\end{equation}
is the spatial projection of the stress-energy tensor, and
\begin{equation}
    S = \gamma^{ij}S_{ij}
\end{equation}
is its trace.

Together with the evolution equation for the spatial metric,
\begin{equation}
    \partial_t \gamma_{ij}
    =
    -2\alpha K_{ij}
    +
    \mathcal{L}_{\beta}\gamma_{ij},
    \label{eq:adm-gamma-evolution}
\end{equation}
these equations form the ADM system. The variables $\gamma_{ij}$ and $K_{ij}$ are the dynamical variables, while $\alpha$ and $\beta^i$ encode the gauge freedom associated with the choice of spacetime coordinates.

\begin{equation}
\boxed{
\begin{aligned}
    R + K^2 - K_{ij}K^{ij}
    &=
    16\pi \rho,
    &&\text{Hamiltonian constraint},
    \\[0.5em]
    D_j \left( K^{ij} - \gamma^{ij}K \right)
    &=
    8\pi j^i,
    &&\text{Momentum constraint},
    \\[0.5em]
    \partial_t \gamma_{ij}
    &=
    -2\alpha K_{ij}
    +
    \mathcal{L}_{\beta}\gamma_{ij},
    &&\text{Metric evolution},
    \\[0.5em]
    \partial_t K_{ij}
    &=
    \alpha
    \left(
        R_{ij}
        - 2K_{ik}K^k{}_j
        + K K_{ij}
    \right)
    - D_i D_j \alpha
    \\
    &\quad
    - 4\pi \alpha
    \left[
        2S_{ij}
        -
        \gamma_{ij}(S-\rho)
    \right]
    +
    \mathcal{L}_{\beta}K_{ij},
    &&\text{Extrinsic-curvature evolution}.
\end{aligned}
}
\end{equation}
\subsection{Reformulating the Einstein Equations}
The ADM equations are actually only weakly hyperbolic, unfortunately they in general have been shown to be numerically unstable. We therefore need to Reformulate them. It has been shown how to reformaulte them into a strongly hyperbolic system of equations. 
\subsubsection{Reformulating Maxwell's equations}
We begin by using the electromagnetism analogy. In a \(3+1\) decomposition,
Maxwell's equations can be written as
\begin{equation}
    \partial_t E_i
    =
    -D_jD^j A_i
    +D_iD^jA_j
    -4\pi j_i ,
\end{equation}
together with
\begin{equation}
    \partial_t A_i
    =
    -E_i-D_i\phi .
\end{equation}
These evolution equations are supplemented by the constraint equation
\begin{equation}
    D_iE^i=4\pi\rho .
\end{equation}
We define the constraint violation \(\mathcal{C}\) by
\begin{equation}
    \mathcal{C}
    \equiv
    D_iE^i-4\pi\rho .
\end{equation}
Ideally, the constraint is satisfied exactly, so that
\begin{equation}
    \mathcal{C}=0 .
\end{equation}
Furthermore, if the constraint is satisfied initially, we would like it to remain
satisfied during the evolution. In exact arithmetic this would imply
\begin{equation}
    \partial_t \mathcal{C}=0 .
\end{equation}
In a numerical evolution, however, truncation error and roundoff error generally
produce small but nonzero values of \(\mathcal{C}\). It is therefore useful to
reformulate the evolution system so that constraint violations propagate in a
controlled way.

We now introduce the auxiliary variable
\begin{equation}
    \Gamma
    \equiv
    D_iA^i .
\end{equation}
Taking a time derivative gives
\begin{equation}
    \partial_t\Gamma
    =
    D_i\partial_t A^i .
\end{equation}
Using the evolution equation for \(A_i\), we find
\begin{equation}
    \partial_t\Gamma
    =
    -D_iE^i-D_iD^i\phi .
\end{equation}
Since \(\mathcal{C}=0\) for constraint-satisfying data, we are free to add an
arbitrary multiple of \(\mathcal{C}\) to the right-hand side. Thus we write
\begin{equation}
    \partial_t\Gamma
    =
    -D_iE^i-D_iD^i\phi+\eta\mathcal{C},
\end{equation}
where \(\eta\) is a constant. Using
\begin{equation}
    \mathcal{C}=D_iE^i-4\pi\rho ,
\end{equation}
this may also be written as
\begin{equation}
    \partial_t\Gamma
    =
    (\eta-1)D_iE^i
    -D_iD^i\phi
    -4\pi\eta\rho .
\end{equation}

With this new variable, the Maxwell evolution system becomes
\begin{align}
    \partial_t A_i
    &=
    -E_i-D_i\phi ,
    \\
    \partial_t E_i
    &=
    -D^2A_i+D_i\Gamma-4\pi j_i ,
    \\
    \partial_t\Gamma
    &=
    -D_iE^i-D_iD^i\phi+\eta\mathcal{C}.
\end{align}
Here \(D^2\equiv D_iD^i\). The advantage of this form is that the constraint
violation no longer simply satisfies \(\partial_t\mathcal{C}=0\). Instead, as we
now show, it satisfies a wave equation.

Starting from
\begin{equation}
    \mathcal{C}
    =
    D_iE^i-4\pi\rho ,
\end{equation}
we take a time derivative:
\begin{equation}
    \partial_t\mathcal{C}
    =
    D_i\partial_t E^i
    -4\pi\partial_t\rho .
\end{equation}
Using charge conservation,
\begin{equation}
    \partial_t\rho
    =
    -D_i j^i ,
\end{equation}
we obtain
\begin{equation}
    \partial_t\mathcal{C}
    =
    D_i\partial_t E^i
    +4\pi D_i j^i .
\end{equation}
Substituting the evolution equation for \(E_i\),
\begin{equation}
    \partial_t E_i
    =
    -D^2A_i+D_i\Gamma-4\pi j_i ,
\end{equation}
gives
\begin{align}
    \partial_t\mathcal{C}
    &=
    D^i\left(-D^2A_i+D_i\Gamma-4\pi j_i\right)
    +4\pi D_i j^i
    \\
    &=
    -D^iD^2A_i+D^iD_i\Gamma .
\end{align}
Assuming that the spatial derivatives commute, this becomes
\begin{equation}
    \partial_t\mathcal{C}
    =
    -D^2D^iA_i+D^2\Gamma .
\end{equation}
Therefore,
\begin{equation}
    \partial_t\mathcal{C}
    =
    D^2\left(\Gamma-D^iA_i\right).
\end{equation}

Taking another time derivative, we find
\begin{equation}
    \partial_t^2\mathcal{C}
    =
    D^2\left(\partial_t\Gamma-D^i\partial_t A_i\right).
\end{equation}
Using
\begin{equation}
    \partial_t A_i
    =
    -E_i-D_i\phi ,
\end{equation}
we have
\begin{equation}
    -D^i\partial_t A_i
    =
    D^iE_i+D^iD_i\phi .
\end{equation}
Hence,
\begin{align}
    \partial_t\Gamma-D^i\partial_t A_i
    &=
    \left(-D^iE_i-D^iD_i\phi+\eta\mathcal{C}\right)
    +
    \left(D^iE_i+D^iD_i\phi\right)
    \\
    &=
    \eta\mathcal{C}.
\end{align}
It follows that
\begin{equation}
    \partial_t^2\mathcal{C}
    =
    D^2\left(\eta\mathcal{C}\right).
\end{equation}
Since \(\eta\) is constant, this reduces to
\begin{equation}
    \partial_t^2\mathcal{C}
    =
    \eta D^2\mathcal{C}.
\end{equation}
Equivalently,
\begin{equation}
    \boxed{
    \left(-\partial_t^2+\eta D_iD^i\right)\mathcal{C}=0
    } .
\end{equation}
Thus the constraint violation \(\mathcal{C}\) satisfies a wave equation. For
\(\eta>0\), constraint violations propagate with speed \(\sqrt{\eta}\). Anayltically we haven't changes the nature of the equations but we have changed the numeric behavior as now the error will propagate a way as it turns out we will need to do something similar for the bssn equations. 

\subsubsection{BSSN Equations}

We begin with the ADM equations:
\begin{align}
    R + K^2 - K_{ij}K^{ij}
    &=
    16\pi \rho,
    &&\text{Hamiltonian constraint},
    \\[0.5em]
    D_j \left( K^{ij} - \gamma^{ij}K \right)
    &=
    8\pi j^i,
    &&\text{Momentum constraint},
    \\[0.5em]
    \partial_t \gamma_{ij}
    &=
    -2\alpha K_{ij}
    +
    \mathcal{L}_{\beta}\gamma_{ij},
    &&\text{Metric evolution},
    \\[0.5em]
    \partial_t K_{ij}
    &=
    \alpha
    \left(
        R_{ij}
        - 2K_{ik}K^k{}_{j}
        + K K_{ij}
    \right)
    - D_i D_j \alpha
    \notag \\
    &\quad
    - 4\pi \alpha
    \left[
        2S_{ij}
        -
        \gamma_{ij}(S-\rho)
    \right]
    +
    \mathcal{L}_{\beta}K_{ij},
    &&\text{Extrinsic-curvature evolution}.
\end{align}

The Ricci tensor can be written in several different forms. In particular,
\(R_{ij}\) contains second derivatives of the metric:
\begin{equation}
    R_{ij}
    =
    -\frac{1}{2}\gamma^{kl}
    \left(
        \partial_k\partial_l\gamma_{ij}
        +
        \partial_i\partial_j\gamma_{kl}
        -
        \partial_i\partial_l\gamma_{kj}
        -
        \partial_k\partial_j\gamma_{il}
    \right)
    +
    \gamma^{kl}
    \left(
        \Gamma^{m}{}_{il}\Gamma_{mkj}
        -
        \Gamma^{m}{}_{ij}\Gamma_{mkl}
    \right).
\end{equation}

The most important terms for rewriting the equations are the four terms involving
second derivatives of the metric. The first term,
\(\gamma^{kl}\partial_k\partial_l\gamma_{ij}\), acts like a Laplace operator
on the metric. However, because of the remaining second-derivative terms, we
cannot immediately write the system in wave-equation form. In analogy with the
introduction of an auxiliary variable in Maxwell's equations, we introduce a
new variable here as well. This allows us to rewrite the Ricci tensor in terms
of the trace of the Christoffel symbols. In particular, we define
\begin{equation}
    \Gamma^i \equiv \gamma^{jk}\Gamma^i{}_{jk}.
\end{equation}
In terms of this contracted Christoffel symbol, the Ricci tensor can be written
as
\begin{equation}
\begin{aligned}
    R_{ij}
    &=
    -\frac{1}{2}\gamma^{kl}\partial_k\partial_l\gamma_{ij}
    +
    \gamma_{k(i}\partial_{j)}\Gamma^k
    +
    \Gamma^k\Gamma_{(ij)k}
    \\
    &\quad
    +
    2\gamma^{lm}\Gamma^k{}_{m(i}\Gamma_{j)kl}
    +
    \gamma^{kl}\Gamma^m{}_{il}\Gamma_{mkj}.
\end{aligned}
\end{equation}

We can also use the identity relating the contracted Christoffel symbols to the
divergence of the inverse metric:
\begin{equation}
    g^{bc}\Gamma^a{}_{bc}
    =
    -\frac{1}{\sqrt{|g|}}
    \partial_b
    \left(
        \sqrt{|g|}g^{ab}
    \right).
\end{equation}
Applying this identity to the spatial metric gives
\begin{equation}
    \label{eq:Gamma_eq}
    \Gamma^i
    =
    -\frac{1}{\sqrt{\gamma}}
    \partial_j
    \left(
        \sqrt{\gamma}\gamma^{ij}
    \right).
\end{equation}

We can now regard \(\Gamma^i\) as an independent variable. It is related to the
divergence of the metric in much the same way as the auxiliary variable in the
Maxwell example above. To complete the system, we must derive an evolution
equation for \(\Gamma^i\). This is done by taking a time derivative of
Eq.~\eqref{eq:Gamma_eq}. Doing so introduces divergences of the extrinsic
curvature. These terms can then be simplified using the momentum constraint.

To avoid some of the difficulties associated with the determinant of the metric,
we also introduce a conformal transformation. In particular, we write
\begin{equation}
    \gamma_{ij}
    =
    e^{4\phi}\tilde{\gamma}_{ij}.
\end{equation}
There is some freedom in the choice of conformal factor. We can then rewrite
quantities such as the Christoffel symbols and the Ricci tensor in terms of the
conformal metric plus correction terms involving the conformal factor. The
physical Ricci tensor can be split into two pieces,
\begin{equation}
    R_{ij}
    =
    \bar{R}_{ij}
    +
    R^{\phi}_{ij},
\end{equation}
where \(\bar{R}_{ij}\) is computed from the conformally related metric
\(\bar{\gamma}_{ij}\), and \(R^{\phi}_{ij}\) contains the terms involving the
conformal factor \(\phi\).

The conformal Ricci tensor is
\begin{equation}
\begin{aligned}
    \bar{R}_{ij}
    &=
    -\frac{1}{2}\bar{\gamma}^{kl}
    \partial_k\partial_l \bar{\gamma}_{ij}
    +
    \bar{\gamma}_{k(i}\partial_{j)}\bar{\Gamma}^{k}
    +
    \bar{\Gamma}^{k}\bar{\Gamma}_{(ij)k}
    \\
    &\quad
    +
    2\bar{\gamma}^{lm}
    \bar{\Gamma}^{k}{}_{m(i}\bar{\Gamma}_{j)kl}
    +
    \bar{\gamma}^{kl}
    \bar{\Gamma}^{m}{}_{il}\bar{\Gamma}_{mkj}.
\end{aligned}
\end{equation}

The conformal-factor contribution is
\begin{equation}
\begin{aligned}
    R^{\phi}_{ij}
    &=
    -2
    \left(
        \bar{D}_i\bar{D}_j\phi
        +
        \bar{\gamma}_{ij}\bar{\gamma}^{lm}
        \bar{D}_l\bar{D}_m\phi
    \right)
    \\
    &\quad
    +
    4
    \left(
        \left(\bar{D}_i\phi\right)
        \left(\bar{D}_j\phi\right)
        -
        \bar{\gamma}_{ij}\bar{\gamma}^{lm}
        \left(\bar{D}_l\phi\right)
        \left(\bar{D}_m\phi\right)
    \right).
\end{aligned}
\end{equation}

The contracted conformal Christoffel symbols are defined by
\begin{equation}
    \bar{\Gamma}^{i}
    \equiv
    \bar{\gamma}^{jk}\bar{\Gamma}^{i}{}_{jk}
    =
    -\frac{1}{\bar{\gamma}^{1/2}}
    \partial_j
    \left(
        \bar{\gamma}^{1/2}
        \bar{\gamma}^{ij}
    \right),
\end{equation}
where \(\bar{\gamma}\) denotes the determinant of the conformal metric
\(\bar{\gamma}_{ij}\).

Using the freedom in the choice of conformal factor, we may impose
\begin{equation}
    \bar{\gamma}=1,
\end{equation}
which reduces the conformal connection functions to
\begin{equation}
    \bar{\Gamma}^{i}
    =
    -\partial_j\bar{\gamma}^{ij}.
\end{equation}

Taking the determinant of
\begin{equation}
    \gamma_{ij}
    =
    e^{4\phi}\bar{\gamma}_{ij},
\end{equation}
and using \(\bar{\gamma}=1\), we obtain
\begin{equation}
    \phi
    =
    \frac{1}{12}\ln\gamma,
\end{equation}
where \(\gamma\) is the determinant of the physical metric \(\gamma_{ij}\).

The extrinsic curvature is decomposed into its trace and trace-free parts:
\begin{equation}
    K_{ij}
    =
    A_{ij}
    +
    \frac{1}{3}\gamma_{ij}K,
\end{equation}
where
\begin{equation}
    K
    =
    \gamma^{ij}K_{ij}.
\end{equation}
Finally, the trace-free part is conformally rescaled according to
\begin{equation}
    A_{ij}
    =
    e^{4\phi}\tilde{A}_{ij}.
\end{equation}

The variables we evolve are therefore
\[
    \phi,\qquad
    \tilde{\gamma}_{ij},\qquad
    \tilde{A}_{ij},\qquad
    K,\qquad
    \tilde{\Gamma}^i.
\]
These are the standard BSSN variables.

\noindent\textbf{The BSSN equations are}
\begin{equation}
\boxed{
\begin{aligned}
    \partial_t \phi
    &=
    -\frac{1}{6}\alpha K
    +
    \beta^i \partial_i \phi
    +
    \frac{1}{6}\partial_i \beta^i ,
    &&\text{Conformal exponent evolution},
    \\[0.75em]
    \partial_t K
    &=
    -\gamma^{ij}D_iD_j\alpha
    +
    \alpha
    \left(
        \tilde{A}_{ij}\tilde{A}^{ij}
        +
        \frac{1}{3}K^2
    \right)
    +
    4\pi\alpha(\rho+S)
    +
    \beta^i\partial_i K ,
    &&\text{Mean-curvature evolution},
    \\[0.75em]
    \partial_t \tilde{\gamma}_{ij}
    &=
    -2\alpha\tilde{A}_{ij}
    +
    \beta^k\partial_k\tilde{\gamma}_{ij}
    +
    \tilde{\gamma}_{ik}\partial_j\beta^k
    +
    \tilde{\gamma}_{kj}\partial_i\beta^k
    -
    \frac{2}{3}\tilde{\gamma}_{ij}\partial_k\beta^k ,
    &&\text{Conformal-metric evolution},
    \\[0.75em]
    \partial_t \tilde{A}_{ij}
    &=
    e^{-4\phi}
    \left[
        -\left(D_iD_j\alpha\right)^{\mathrm{TF}}
        +
        \alpha
        \left(
            R_{ij}^{\mathrm{TF}}
            -
            8\pi S_{ij}^{\mathrm{TF}}
        \right)
    \right]
    +
    \alpha
    \left(
        K\tilde{A}_{ij}
        -
        2\tilde{A}_{il}\tilde{A}^{l}{}_{j}
    \right)
    \\
    &\quad
    +
    \beta^k\partial_k\tilde{A}_{ij}
    +
    \tilde{A}_{ik}\partial_j\beta^k
    +
    \tilde{A}_{kj}\partial_i\beta^k
    -
    \frac{2}{3}\tilde{A}_{ij}\partial_k\beta^k ,
    &&\text{Trace-free curvature evolution},
    \\[0.75em]
    \partial_t \tilde{\Gamma}^{i}
    &=
    -2\tilde{A}^{ij}\partial_j\alpha
    +
    2\alpha
    \left(
        \tilde{\Gamma}^{i}{}_{jk}\tilde{A}^{kj}
        -
        \frac{2}{3}\tilde{\gamma}^{ij}\partial_j K
        -
        8\pi\tilde{\gamma}^{ij}j_j
        +
        6\tilde{A}^{ij}\partial_j\phi
    \right)
    \\
    &\quad
    +
    \beta^j\partial_j\tilde{\Gamma}^{i}
    -
    \tilde{\Gamma}^{j}\partial_j\beta^i
    +
    \frac{2}{3}\tilde{\Gamma}^{i}\partial_j\beta^j
    +
    \frac{1}{3}\tilde{\gamma}^{li}\partial_l\partial_j\beta^j
    +
    \tilde{\gamma}^{lj}\partial_j\partial_l\beta^i ,
    &&\text{Conformal-connection evolution}.
\end{aligned}
}
\end{equation}

\subsubsection{Gauge and Slicing Conditions}

We are not quite done, because we still have freedom to choose our coordinate
system. In particular, we must choose conditions for the lapse \(\alpha\) and
the shift \(\beta^i\). We summarize several common choices below.

\paragraph{Geodesic slicing.}
A simple choice is to set
\begin{equation}
    \alpha=1,
    \qquad
    \beta^i=0.
\end{equation}
This is called geodesic slicing. Although this is an easy choice, it is usually
a poor one for numerical relativity simulations. The acceleration of the normal
observers is
\begin{equation}
    a_a
    =
    D_a\ln\alpha .
\end{equation}
For geodesic slicing, \(\alpha=1\), so \(a_a=0\). Thus the normal observers are
in free fall. This is also true for the coordinate points. In a strong
gravitational field, this can lead to coordinate singularities, making geodesic
slicing unsuitable for most numerical simulations of black holes.

\paragraph{Maximal slicing.}
We can control the convergence of coordinate observers by controlling the mean
curvature. In maximal slicing, we impose
\begin{equation}
    K=0.
\end{equation}
We have imposed this condition on initial data in the past, but here we impose
it at all times. Therefore,
\begin{equation}
    \partial_t K=0.
\end{equation}
This condition has several desirable properties, including singularity avoidance.
In particular, the lapse tends to collapse toward zero near strong-field
regions. This causes those slices to be held back in the evolution, a phenomenon
known as the collapse of the lapse. However, maximal slicing also has
disadvantages, such as grid stretching, which can lead to sharp features that
are difficult to resolve numerically.

\paragraph{Harmonic coordinates.}
Harmonic coordinates are defined by the condition
\begin{equation}
    \Gamma^a=0.
\end{equation}
A related condition is harmonic slicing, in which we require only the time
component of \(\Gamma^a\) to vanish and set \(\beta^i=0\). To obtain a lapse
condition, we compute
\begin{equation}
    \Gamma^0
    =
    -\frac{1}{\alpha\gamma^{1/2}}
    \partial_t
    \left(
        \gamma^{1/2}\alpha^{-1}
    \right)
    =
    0.
\end{equation}
This gives the lapse evolution equation
\begin{equation}
    \partial_t \alpha
    =
    -\alpha^2 K.
\end{equation}
Solutions to the resulting wave equation are generally well behaved. However,
with a few modifications to harmonic slicing, we can obtain an even better gauge
choice for black-hole evolutions.

\paragraph{\(1+\log\) slicing and the Gamma-driver condition.}
We generalize the lapse evolution equation to
\begin{equation}
    \partial_t \alpha
    =
    -\alpha^2 f(\alpha)K.
\end{equation}
For \(f(\alpha)=1\), this reduces to harmonic slicing. For
\(f(\alpha)=0\), the lapse is constant in time, which gives geodesic slicing
when \(\alpha=1\). In the limit \(f(\alpha)\rightarrow\infty\), this condition
approaches maximal slicing.

A particularly useful choice is
\begin{equation}
    f(\alpha)
    =
    \frac{2}{\alpha}.
\end{equation}
This gives
\begin{equation}
    \partial_t\alpha
    =
    -2\alpha K,
\end{equation}
which is the standard \(1+\log\) slicing condition. This choice has the
singularity-avoidance properties of maximal slicing while retaining good
behavior in nearly flat regions.

As a further generalization, we do not assume that \(\beta^i\) is close to zero.
We include the advective term and write
\begin{equation}
    \left(\partial_t-\beta^i\partial_i\right)\alpha
    =
    -\alpha^2 f(\alpha)K.
\end{equation}
For \(f(\alpha)=2/\alpha\), this becomes
\begin{equation}
    \left(\partial_t-\beta^i\partial_i\right)\alpha
    =
    -2\alpha K.
\end{equation}

We also need an evolution condition for the shift. A common choice is the
Gamma-driver condition,
\begin{equation}
    \left(\partial_t+\beta^j\partial_j\right)\beta^i
    =
    \mu B^i,
\end{equation}
together with
\begin{equation}
    \left(\partial_t+\beta^j\partial_j\right)B^i
    =
    \left(\partial_t+\beta^j\partial_j\right)\tilde{\Gamma}^i
    -
    \eta B^i.
\end{equation}
Here \(\mu\) and \(\eta\) are constants. The combination of \(1+\log\) slicing
and the Gamma-driver shift condition is called the moving-puncture gauge. A
major advantage of this gauge is that it removes the need for excision: the
singularities are effectively handled by the gauge itself.


\section{Equations of Motion}
\label{sec:appx:eom}

The action for EMDA in the Einstein frame is given as:
\begin{equation}
\begin{aligned}
S=\int \mathrm{d}^4 x\,\sqrt{-g}\Big[
&{R\over8\pi}-2(\nabla\phi)^2-e^{-2\alpha_0\phi}F^2\\
&-\tfrac12 e^{4\alpha_1\phi}(\nabla\kappa)^2
-\kappa\,F_{ab}(*F)^{ab}
\Big]
\end{aligned}
\end{equation}
where $\phi$ and $\kappa$ are the dilaton and axion, respectively.
The tensor $F_{ab}$ is the U(1) field strength, and $(*F)_{ab}$ is its
dual. In this work we set the coupling constants to
$\alpha_0=\alpha_1=1$, corresponding to a low-energy limit of heterotic
string theory.

The Einstein equations are:
\begin{equation}
\begin{split}
%R_{ab} - \tfrac{1}{2} g_{ab} R
G_{ab} = 8\pi G\bigg[&
 \, 2\left(
    \nabla_a \phi \nabla_b \phi
    - \tfrac{1}{2} g_{ab}(\nabla \phi)^2
\right) \\
&+ 2 e^{-2\alpha_0\phi}
\left(
    F_{ac}F_b{}^c
    - \tfrac{1}{4} g_{ab}F^2
\right) \\
&+ \tfrac{1}{2} e^{4\alpha_1\phi}
\left(
    \nabla_a\kappa \nabla_b\kappa
    - \tfrac{1}{2} g_{ab}(\nabla \kappa)^2
\right)
\bigg] 
\end{split}
\label{eq:Einstein}
\end{equation}
while the matter equations can be written 
\begin{align}
0 &= \nabla^a \nabla_a \phi
   + \tfrac{1}{2}\alpha_0 e^{-2\alpha_0 \phi} F^2
   - \tfrac{1}{2}\alpha_1 e^{4\alpha_1 \phi} (\nabla \kappa)^2, \label{eq:phi} \\[1ex]
0 &= \nabla^a \!\left( e^{4\alpha_1 \phi} \nabla_a \kappa \right)
   - F_{ab} (\ast F)^{ab}, \label{eq:kappa} \\[1ex]
0 &= \nabla_a \!\left( e^{-2\alpha_0 \phi} F^{ab} \right)
   + (\ast F)^{ab} \nabla_a \kappa. \label{eq:Maxwell}
\end{align}

We include a form of hyperbolic constraint damping into the Maxwell equations so that violations propagate away at the speed of light. In particular, we introduce two scalar fields $\Phi$ and $\Psi$ into \ref{eq:Maxwell} and the identity $\nabla_a(*F)^{ab}=0$ as follows:
\begin{align}
\nabla_{a}\left(F^{ab} + g^{ab}\Psi \right) 
  &= \eta_{1} n^{a} \Psi 
     + 2 \alpha_{0} \nabla_{a}\phi\, F^{ab} \nonumber \\
  &\quad - e^{2\alpha_0 \phi} (*F)^{ab}\nabla_a \kappa, \\[6pt]
\nabla_{a}\!\left( (\ast F)^{ab} + g^{ab}\Phi \right) 
  &= \eta_{2} n^{a} \Phi .
\end{align}
where $\eta_1$ and $\eta_2$ are taken as positive constants and serve as tunable parameters to effect the damping of the elctromagnetic constraints.  Note that on the right hand sides, we have also included $n^a$, the usual unit timelike normal to the spatial hypersurfaces in anticipation of performing the usual $3+1$ split of spacetime.  It is worth noting that these damped versions of Maxwell are now no longer strictly 4-covariant.  

We introduce the usual $3+1$ decomposition of spacetime by defining the line element as 
\begin{align}
{\rm d}s^2 = -\alpha^2 {\rm d} t^2 + \gamma_{ij} ({\rm d}x^i + \beta^i {\rm d}t) ({\rm d}x^j + \beta^j {\rm d}t) 
\end{align}
with lapse, $\alpha$, shift, $\beta^i$, 3-metric, $\gamma_{ij}$, unit timelike normal $n_a = (-\alpha,0,0,0)$ and extrinsic curvature $K_{ab} = - \gamma_a{}^c \gamma_b{}^d\nabla_c n_d$.  

Using the BSSN formulation, we solve the standard 
BSSN evolution equations for the usual $\chi, {\tilde \gamma}_{ij}, K, {\tilde A}_{ij}$ and ${\tilde \Gamma}^i$ variables with EMDA matter terms.\footnote{Note that in this Appendix only, $\chi$ represents the conformal factor and not the dimensionless spin parameter.}  These equations take the form 

\begin{align}
\partial_0\tilde{\gamma}_{ij} &= -2\alpha\,\tilde{A}_{ij},\\[4pt]
\partial_0 \chi &= \tfrac{2}{3}\alpha\,\chi\,K,\\[4pt]
\partial_0 K &= [\ldots] + 4\pi\alpha\,(\rho + T),\\[4pt]
\partial_0\tilde{A}_{ij} &= [\ldots] - 8\pi\chi\alpha\left( T_{ij} - \tfrac{1}{3}\,\gamma_{ij} T_k{}^k\right),\\[4pt]
\partial_0 \tilde{\Gamma}^i &= [\ldots] - 16\pi\alpha\,\chi^{-1} S^{\,i},
%\tilde{\Gamma}^i &= \tilde{\gamma}^{jk}\tilde{\Gamma}^{\,i}{}_{jk}
\end{align}
where $\partial_0 \equiv \partial_t - {\cal L}_{\beta}$ with ${\cal L}_{\beta}$ the Lie derivative along the shift and the $[\ldots]$ are a stand-in for the relevant right hand sides which can be found in standard references, e.g.~\cite{Baumgarte2010} 

The quantities $S^i$, $T_{ij}$, and $\rho$ are defined as
\begin{equation}
\begin{aligned}
S^i
&\equiv -\gamma^{ia} T_{ab} n^b  \\
&=   2\Pi D^i\phi
   + 2e^{-2\alpha_0\phi}\epsilon^{ijk}E_jB_k  %\\
%&\qquad\qquad
   + \tfrac{1}{2}e^{4\alpha_1\phi}\Xi D^i\kappa ,
\label{eq:Si}
\end{aligned}
\end{equation}
\begin{equation}
\begin{aligned}
T_{ij}
&\equiv T_{ab} \gamma^a{}_i \gamma^b{}_j  \\
&= 2\Bigl[
      D_i\phi D_j\phi
      - \tfrac{1}{2}\tilde{\gamma}_{ij}\bigl((D\phi)^2-\Pi^2\bigr)
   \Bigr]  \\
&\quad\quad
- 2e^{-2\alpha_0\phi}\Bigl( 
       E_iE_j + B_iB_j  \\
&\qquad\qquad\qquad\quad
      - \tfrac{1}{2}\tilde{\gamma}_{ij}
        \bigl(E_kE^k + B_kB^k\bigr)
   \Bigr)  \\
&\quad\,
+ \tfrac{1}{2}e^{4\alpha_1\phi}\Bigl[
      D_i\kappa D_j\kappa
      - \tfrac{1}{2}\tilde{\gamma}_{ij}\bigl((D\kappa)^2-\Xi^2\bigr)
   \Bigr] .
\label{eq:Tij}
\end{aligned}
\end{equation}

\begin{equation}
\begin{split}
    \rho \equiv\;& \Pi^2 + (D\phi)^2 
    + e^{-2 \alpha_0 \phi}\big( E_a  E^a +  B_a  B^a\big) \\
    &+ \tfrac{1}{4} e^{4 \alpha_1 \phi}\big(\Xi^2 + (D\kappa)^2\big).
\end{split}
\end{equation}

We use standard puncture gauge conditions for the lapse and the shift:
\begin{align}
\partial_t\alpha &=\beta^i \partial_i - 2\alpha K , \\[4pt]
\partial_t \beta^i &= \beta^j \partial_j \beta^i + \tfrac{3}{4} B^i , \\[4pt] 
\partial_t B^i &= \beta^j \partial_j B^i + \partial_t {\tilde\Gamma}^i - \beta^j \partial_j {\tilde \Gamma}^i - \eta B^i .
\end{align}
We also employ a slow-start lapse condition~\cite{Etienne2024} to reduce transients at early times.  This amounts to adding the term 
\begin{equation}
- \chi^{1/2} \bigl(\varpi \, e^{-t^2/2\sigma^2}\bigr)\,(\alpha - \chi^{1/2}) \\[4pt]
\end{equation}
to the equation for the lapse.  Following~\cite{Etienne2024}, we take the constants $\varpi=3/5$ and $\sigma=20$.

For the matter equations we start with the dilaton and axion:
\begin{align}
    \partial_0 \phi &= - \alpha \Pi , \\
    \partial_0 \kappa &= - \Xi 
\end{align}
\begin{equation}
\begin{aligned}
\partial_0 \Pi
=& \, \alpha K \Pi
- \alpha \chi\, \tilde{\gamma}^{ij} \partial_i \partial_j \phi
- \chi\, \tilde{\gamma}^{ij} \partial_i \alpha\, \partial_j \phi  \\
&+ \alpha \chi\, \tilde{\Gamma}^i \partial_i \phi
+ \frac{\alpha}{2}\tilde{\gamma}^{ij}
  \partial_i \chi\, \partial_j \phi \\
&+ \frac{\alpha\alpha_0}{\chi}
  e^{-2\alpha_0\phi}
  \tilde{\gamma}_{ij}
  \left(E^iE^j - B^iB^j\right)  \\
&+ \frac{\alpha\alpha_1}{2}
  e^{4\alpha_1\phi}
  \left(
    \chi\,\tilde{\gamma}^{ij}
    \partial_i\kappa\,\partial_j\kappa
    - \Xi^2
  \right).
\end{aligned}
\end{equation}


\begin{equation}
\begin{aligned}
\partial_0 \Xi
&= \alpha K \Xi
\\[3pt]
&\quad
- \alpha \chi\, \tilde{\gamma}^{ij} \partial_i \partial_j \kappa
- \chi\, \tilde{\gamma}^{ij}\, \partial_i \alpha\, \partial_j \kappa
\\[3pt]
&\quad
+ \alpha \chi\, \tilde{\Gamma}^i \partial_i \kappa
+ \frac{\alpha}{2}\, \tilde{\gamma}^{ij}\, \partial_i \kappa\, \partial_j \chi
\\[6pt]
&\quad
- 4 \alpha \alpha_1 \chi\, \tilde{\gamma}^{ij}\,
  \partial_i \kappa\, \partial_j \phi
+ 4 \alpha \alpha_1 \Pi \Xi
\\[6pt]
&\quad
+ \frac{4\alpha}{\chi}\,
  e^{-4\alpha_1 \phi}\,
  \tilde{\gamma}_{ij} B^i E^j .
\end{aligned}
\end{equation}
The equations for the electromgnetic fields, including hyperbolic divergence cleaning become
\begin{equation}
\begin{aligned}
\partial_0 \Psi
&= - \alpha\, \partial_i E^i
\\[3pt]
&\quad
+ \frac{3\alpha}{2\chi}\, E^i \partial_i \chi
+ 2 \alpha_0 \alpha\, E^i \partial_i \phi
\\[3pt]
&\quad
+ \alpha e^{2\alpha_0 \phi}\, B^i \partial_i \kappa
- \alpha \eta_1 \Psi .
\end{aligned}
\end{equation}
\begin{equation}
\begin{aligned}
\partial_0 \Phi
&= \alpha\, \partial_i B^i
- \frac{3\alpha}{2\chi}\, B^i \partial_i \chi
- \alpha \eta_2 \Phi .
\end{aligned}
\end{equation}
\begin{equation}
\begin{aligned}
\partial_0 E^i
&= \alpha K E^i
\\[4pt]
&\quad
+ 2 \alpha \alpha_0 \chi^{1/2}\,
  \epsilon^{jik}\tilde{\gamma}_{kl} B^l \partial_j \phi
- 2 \alpha \alpha_0 E^i \Pi
\\[6pt]
&\quad
+ \alpha e^{2\alpha_0 \phi} \chi^{1/2}\,
  \epsilon^{ijk}\tilde{\gamma}_{kl} E^l \partial_j \kappa
- \alpha e^{2\alpha_0 \phi} \Xi B^i
\\[6pt]
&\quad
- \alpha \chi\, \tilde{\gamma}^{ij} \partial_j \Psi
- \chi^{1/2}\, \epsilon^{jik}\tilde{\gamma}_{kl} B^l \partial_j \alpha
\\[4pt]
&\quad
- \alpha \chi^{1/2}\, \epsilon^{jik} B^l \partial_j \tilde{\gamma}_{kl}
- \alpha \chi^{1/2}\, \epsilon^{jik}\tilde{\gamma}_{kl} \partial_j B^l
\\[4pt]
&\quad
+ \alpha \chi^{-1/2}\, \epsilon^{jik}\tilde{\gamma}_{kl} B^l \partial_j \chi .
\end{aligned}
\end{equation}

\begin{equation}
\begin{aligned}
\partial_0 B^i
&= \alpha K B^i
+ \alpha \chi\, \tilde{\gamma}^{ij} \partial_j \Phi
\\[6pt]
&\quad
- \chi^{1/2}\, \epsilon^{ijk}\tilde{\gamma}_{kl} E^l \partial_j \alpha
+ \alpha \chi^{-1/2}\, \epsilon^{ijk}\tilde{\gamma}_{kl} E^l \partial_j \chi
\\[4pt]
&\quad
- \alpha \chi^{1/2}\, \epsilon^{ijk} E^l \partial_j \tilde{\gamma}_{kl}
- \alpha \chi^{1/2}\, \epsilon^{ijk}\tilde{\gamma}_{kl} \partial_j E^l .
\end{aligned}
\end{equation}